Cycle Route Uptake and Scenario Estimation (CRUSE): an approach for developing strategic cycle network planning tools

### 2024-08-06 paper.qmd
Add hyperlink to abstract
Update with smaller Figure 2
KableExtra

---
title: "Cycle Route Uptake and Scenario Estimation (CRUSE): an approach for developing strategic cycle network planning tools"
# subtitle: A case study of Ireland
bibliography: references.bib
csl: apa.csl
editor:
  markdown:
    wrap: sentence
execute: 
  echo: false
  message: false
number-sections: true
latex-clean: false
pdf-engine: pdflatex
format: 
  # html: default
  # Uncomment for formatted version
  elsevier-pdf:
    linestretch: 2
    keep-tex: true
    journal:
      name: European Transport Research Review
      formatting: preprint
      model: 3p
      cite-style: super
# # # #   # pdf: default
# # Uncomment the following line to produce the cover page
# author:
#   - name: Robin Lovelace
#     email: r.lovelace@leeds.ac.uk
#     corresponding: true
#     affiliations:
#       - name: University of Leeds
#         address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
#   - name: Joey Talbot
#     affiliations:
#       - name: University of Leeds
#         address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
#   - name: Eugeni Vidal-Tortosa
#     affiliations:
#       - name: University of Leeds
#         address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
#   - name: Hussein Mahfouz
#     affiliations:
#       - name: University of Leeds
#         address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
#   - name: Elaine Brick
#     affiliations:
#       - name: AECOM
#         address: Unit 6, Galway Technology Park, Parkmore, Galway, Ireland
#   - name: Peter Wright
#     affiliations:
#       - name: AECOM
#         address: Winslade House, Winslade Park, Manor Drive, Clyst St Mary, Exeter, EX5 1FY, UK
#   - name: "Gary O'Toole"
#     affiliations:
#       - name: AECOM
#         address: Unit 6, Galway Technology Park, Parkmore, Galway, Ireland
#   - name: Dan Brennan
#     affiliations:
#       - name: Transport Infrastructure Ireland
#         address: Parkgate Business Centre, Parkgate Street, Dublin 8, D08 DK10, Ireland
#   - name: Suzanne Meade
#     affiliations:
#       - name: Transport Infrastructure Ireland
#         address: Parkgate Business Centre, Parkgate Street, Dublin 8, D08 DK10, Ireland
abstract: |
  This paper describes an approach for developing strategic cycle network planning tools. Based on our experience developing and deploying the Cycle Route Uptake and Scenario Estimation (CRUSE) Tool for Ireland, we outline the underlying methods, including disaggregation of origin-destination data with the open source 'odjitter' software, incorporation of additional trip purposes, routing, scenario generation, and development of an intuitive user interface that is tested and used by practitioners. Commissioned by the national infrastructure agency Transport Infrastructure Ireland, CRUSE provides estimates of current and potential future cycling levels under 'snapshot' scenarios to inform investment decisions. The publicly available results at https://cruse.bike/ enable planners, engineers, and other stakeholders to make more evidence-based decisions. CRUSE goes beyond previous work by: modeling networks at high spatial resolution; simulating multiple trip purposes (social, shopping, personal utility, recreational, and cycle touring), supplementing official origin-destination datasets on travel for work and education; and providing estimates of 'quietness' (a proxy for cyclist comfort and route preference) at the route segment level. Three network types --- 'Fastest', 'Balanced', and 'Quietest' --- help plan both arterial and residential cycle networks. Workshops with stakeholders were used to inform the development of the tool. Feedback shows that the tool has a wide range of uses and is already being used in practice to inform urban, inter-urban, and rural cycle network designs. The approach is flexible and open-source, allowing the underlying ideas and code to be adapted, supporting more evidence-based and effective cycling policies and interventions internationally.
keywords: 
  - Cycling
  - Open-source
  - Road Safety
  - Active Travel
  - Transport Planning
  - Collaborative Planning
date: last-modified
---

@@

![](scenarios_kildare.png)

>>>>

---
title: "Cycle Route Uptake and Scenario Estimation (CRUSE): an approach for developing strategic cycle network planning tools"
# subtitle: A case study of Ireland
bibliography: references.bib
csl: apa.csl
editor:
  markdown:
    wrap: sentence
execute: 
  echo: false
  message: false
number-sections: true
# # # #   # pdf: default
# # Uncomment the following line to produce the cover page
author:
  - name: Robin Lovelace
    email: r.lovelace@leeds.ac.uk
    corresponding: true
    affiliations:
      - name: University of Leeds
        address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
  - name: Joey Talbot
    affiliations:
      - name: University of Leeds
        address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
  - name: Eugeni Vidal-Tortosa
    affiliations:
      - name: University of Leeds
        address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
  - name: Hussein Mahfouz
    affiliations:
      - name: University of Leeds
        address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
  - name: Elaine Brick
    affiliations:
      - name: AECOM
        address: Unit 6, Galway Technology Park, Parkmore, Galway, Ireland
  - name: Peter Wright
    affiliations:
      - name: AECOM
        address: Winslade House, Winslade Park, Manor Drive, Clyst St Mary, Exeter, EX5 1FY, UK
  - name: "Gary O'Toole"
    affiliations:
      - name: AECOM
        address: Unit 6, Galway Technology Park, Parkmore, Galway, Ireland
  - name: Dan Brennan
    affiliations:
      - name: Transport Infrastructure Ireland
        address: Parkgate Business Centre, Parkgate Street, Dublin 8, D08 DK10, Ireland
  - name: Suzanne Meade
    affiliations:
      - name: Transport Infrastructure Ireland
        address: Parkgate Business Centre, Parkgate Street, Dublin 8, D08 DK10, Ireland
abstract: |
  This paper describes an approach for developing strategic cycle network planning tools. Based on our experience developing and deploying the Cycle Route Uptake and Scenario Estimation (CRUSE) Tool for Ireland, we outline the underlying methods, including disaggregation of origin-destination data with the open source 'odjitter' software, incorporation of additional trip purposes, routing, scenario generation, and development of an intuitive user interface that is tested and used by practitioners. Commissioned by the national infrastructure agency Transport Infrastructure Ireland, CRUSE provides estimates of current and potential future cycling levels under 'snapshot' scenarios to inform investment decisions. The publicly available results at <https://cruse.bike/> enable planners, engineers, and other stakeholders to make more evidence-based decisions. CRUSE goes beyond previous work by: modeling networks at high spatial resolution; simulating multiple trip purposes (social, shopping, personal utility, recreational, and cycle touring), supplementing official origin-destination datasets on travel for work and education; and providing estimates of 'quietness' (a proxy for cyclist comfort and route preference) at the route segment level. Three network types --- 'Fastest', 'Balanced', and 'Quietest' --- help plan both arterial and residential cycle networks. Workshops with stakeholders were used to inform the development of the tool. Feedback shows that the tool has a wide range of uses and is already being used in practice to inform urban, inter-urban, and rural cycle network designs. The approach is flexible and open-source, allowing the underlying ideas and code to be adapted, supporting more evidence-based and effective cycling policies and interventions internationally.
keywords: 
  - Cycling
  - Open-source
  - Road Safety
  - Active Travel
  - Transport Planning
  - Collaborative Planning
date: last-modified
---

```{r}
#| include: false
#| eval: false
latex-clean: false
pdf-engine: pdflatex
format: 
  # html: default
  # Uncomment for formatted version
  elsevier-pdf:
    linestretch: 2
    keep-tex: true
    journal:
      name: European Transport Research Review
      formatting: preprint
      model: 3p
      cite-style: super
```

@@

![](scenarios_kildare.png){width="60%"}

### 2024-08-06 paper.tex
Update with smaller Figure 2
KableExtra

\usepackage{amsmath,amssymb}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\journal{European Transport Research Review}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num}
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Cycle Route Uptake and Scenario Estimation (CRUSE): an approach for developing strategic cycle network planning tools},
  pdfkeywords={Cycling, Open-source, Road Safety, Active
Travel, Transport Planning, Collaborative Planning},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\includegraphics{scenarios_kildare.png}

>>>>

\usepackage{amsmath,amssymb}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

@@

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\journal{European Transport Research Review}

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num}
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Cycle Route Uptake and Scenario Estimation (CRUSE): an approach for developing strategic cycle network planning tools},
  pdfauthor={Robin Lovelace; Joey Talbot; Eugeni Vidal-Tortosa; Hussein Mahfouz; Elaine Brick; Peter Wright; Gary O'Toole; Dan Brennan; Suzanne Meade},
  pdfkeywords={Cycling, Open-source, Road Safety, Active
Travel, Transport Planning, Collaborative Planning},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\includegraphics[width=0.6\textwidth,height=\textheight]{scenarios_kildare.png}
